SNEQ_Lynott_example

### 2017-10-18 SNEQ_Lynott_example.Rmd

Lynott et al. (2014), replicating Williams & Bargh (2008) conducted a study to investigate the effect of being exposed to physical warmth on interpersonal warmth and prosocial behavior. They observed  that 51.2% of participants who received a cold pack (n = 427) opted to receive a reward for themselves, while 57.9% (n = 434) of participants who received a warm pack did the same. Calculating a z-test for the difference between the proportions, this effect is statistically significant (Diff = -0.067, z = -1.98, p = 0.024). Having obtained a significant result, we are left with two alternatives:

1. The difference is small enough that even though it is *statisitcally* significant, we can consider it *practically* equivalent to zero.
2. The difference is substantial enough that we cannot rule out that it is also practically significant.

@@

Using the TOSTtwo.prop functiion in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Lynott et al. (2014), assuming an aplha of 0.05, and using +/- the value of `crit.diff` (`r crit.diff`) as our equivalence bounds. We find that the TOST is non-significant, t(96.62) = 0.72, p = 0.24. Thus, in addition to begin statistically different from zero, we cannot reject the possibility that the true difference really is larger than our SESOI.

>>>>

Lynott et al. (2014) conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on ubsequent judgments related to interpersonal warmth and prosocial behavior (replicating Williams and Bargh, 2008). They observed that 51.2% of participants who received a cold pack (n = 427) opted to receive a reward for themselves, while 57.9% (n = 434) of participants who received a warm pack did the same. Calculating a z-test for the difference between the proportions, this effect is statistically significant (Diff = -0.067, z = -1.98, p = 0.024). Having obtained a significant result, we are left with two alternatives:

1. The difference is small enough that even though it is *statistically* significant, we can consider it *practically* equivalent to zero.
2. The difference is substantial enough that we cannot rule out that it is also practically significant.

@@

Using the TOSTtwo.prop function in the TOSTER package in R, we can then calculate the TOST by insterting the summary statistics from Lynott et al. (2014), assuming an aplha of 0.05, and using +/- the value of `crit.diff` (`r crit.diff`) as our equivalence bounds. We find that the TOST is non-significant, t(96.62) = 0.72, p = 0.24. Thus, in addition to begin statistically different from zero, we cannot reject the possibility that the true difference really is larger than our SESOI.

### 2017-10-18 manuscript.Rmd

An equivalence test can be used to test whether the observed effect is surprisingly close to zero, assuming a meaningful effect exists in the population. The test is a simple variation on the widely used null hypothesis significance tests, with the small difference that instead of testing the observed effect against zero, it is tested against a smallest effect size of interest (SESOI). For example, if a difference on a math test of 5% or more is considered meaningful, one could test against a SESOI of 5%. If the observed effect is suprisingly closer to 0, assuming a true effect of 5% or larger existed, one can declare statistical equivalence, or reject the presence of effects that are large enough to be considered meaningful. An equivalence test can be used to falsify hypotheses.

One can conclude statistical equivalence whenever a 90% confidence interval around the effect size estimate does not contain the SESOI, or when two one-sided tests (i.e., examining whether the difference on the math test is more than -5%, but less than 5%) are statistically significant. By combining null-hypothesis tests and equivalence tests, four combinations of results can be observed (see Figure 1). Two outcomes are straightforward to interpret, namely when the effect is equivalent to zero, and not different from zero (an effect that is not meaningful), and when the effect is not equivalent, and different from zero (a meaningful effect). In addition, the effect can be different from zero and too small to be considered meaningful (an effect which is not meaningful) or the effect can neither differ from zero, not be equivalent to zero (an undetermined result, where more data is needed).

Even though equivalence tests are just a small variation of traditional Frequentist tests (testing against the SESOI, instead of 0), their use was limited until the availability of user-friendly software to perform the calculations [@lakens_equivalence_2017]. In this article, we discuss different approaches to determining the SESOI for psychological research, and provide detailed reproducible examples of how to perform power analyses when designing equivalence tests, and statistical re-analyses of published psychology experiments.

<!-- If we use Dienes & McLatchie's 5 examples, it would make sense to add a Figure like a forest plot, where we present the effect size estimates (raw or standardized) indicated by a square, lines that represent the 90% CI, and another symbol or | to indicate the SESOI, sor all examples. We can then also use the same Figure in the intro to explain the four possible outcomes of equivalence testing). Although we should do all tests in R, we also should try to create Jamovi files that contain the data and analyses - another reason to follow Zoltan in using the Social Psychology special issue studies, because the data should be shared, and we can import the raw data into Jamovi files, save these (with the analysis) and share them.  -->

@@

Equivalence tests are performed against a value that is considered the smallest effect size of interest (SESOI). The SESOI can sometimes based on just noticeable differences, which can be objectively determined. Most often, however, it is a subjective decision that varies across individuals and time. Both these approaches are valid, and even when a just noticeable difference can be objectively determined, researchers might choose to set the SESOI to a larger value. In many research areas, the SESOI is best based on a cost-benefit analysis. Since both costs and benefits are necessarily relative, the SESOI will depend on the researcher who designs the study. The goal of setting a SESOI is to clearly justify why designing a study that has a high probability of rejecting effects larger than a specified value contributes to our knowledge base. Researchers should not aim to determine a SESOI that is universally valid. The goal is to set a SESOI such that inferences based on it answer a meaningful question.

@@

>>>>

An equivalence test can be used to test whether the observed effect is surprisingly small, assuming a meaningful effect exists in the population. The test is a simple variation on the widely used null hypothesis significance tests. To understand the idea behind equivalence tests, it is useful to realize the null hypothesis we test against can be any numerical value. When we compare two groups, we often test whether the difference between groups is zero.

Imagine a researcher who who is interested in voluntary participation in a national program to train young infants' motor skills. The researcher wants to test whether more boys than girls are brought in to the training program by their parents. One could test whether the difference betwen boys and girls is zero. However, this would ignore that the human sex ratio of boys to girls is not exactly 1:1, or an odds ratio of 1. On average, slightly more boys than girls are born. Given that a null hypothesis of exactly zero is unlikely, the researcher tests whether the difference between boys and girls is surprising, assuming an odds ratio in the range of 0.93 to 1.07. If the null hypothesis can be rejected (indicating the effect is surprisingly smaller than OR 0.93, or sirprisingly larger than 1.07), the researcher decides to act as if there is a real gender difference in participation in the motor skill training program. This test against is known as a *minimal effect test*.

However, even if the researcher would consider any odds ratio statistically smaller than 0.93 or larger than 1.07 a rejection of the null hypothesis, not every difference is large enough to matter.  After an extensive discussion with experts, the researcher decides that any effect smaller than an odds ratio of 0.8 or larger than an odds ratio of 1.2 is substantial enough to move money away from the core training program, and address the gender imbalance in the voluntary participation rate. To test whether the gender difference is not large enough to matter, the researcher can perform an *equivalence test*, which tests whether the observed gender difference is statistically larger than an odds ratio of 0.8, *and* statistically smaller than an odds ratio of 1.2. If this is the case, the researcher can act as if the gender difference is too small to intervene.

One can conclude statistical equivalence whenever a 90% confidence interval around the observed estimate does not contain the smallest effect size of interest (SESOI; the values 0.8 and 1.2 in the previous example). When two one-sided tests (i.e., one testing whether the odds ratio is larger than 0.8 the other testing whether the odds ration is smaller than 1.2) are statistically significant, we can conclude statistical equivalence. By combining null-hypothesis tests and equivalence tests, four combinations of results can be observed (see Figure 1). Two outcomes are straightforward to interpret, namely when the effect is equivalent to zero, and not different from zero (an effect that is not meaningful), and when the effect is not equivalent, and different from zero (a meaningful effect). In addition, the effect can be different from zero and too small to be considered meaningful (an effect which is not meaningful) or the effect can neither differ from zero, not be equivalent to zero (an undetermined result, where more data is needed).

Even though equivalence tests are just a small variation of traditional frequentist null-hypothesis significance tests (and test against the SESOI, instead of 0), their use was limited until the availability of user-friendly software to perform the calculations [@lakens_equivalence_2017]. In this article, we provide several examples of equivalence tests which illustrate the procedure. Furthermore, we discuss different approaches to determining the SESOI for psychological research, and provide detailed reproducible examples of how to perform power analyses when designing equivalence tests, and statistical re-analyses of published psychology experiments.

@@

Equivalence tests are performed against a value that is considered the smallest effect size of interest (SESOI). The SESOI can sometimes based on just noticeable differences, which can be objectively determined. Most often, however, it is a subjective decision that varies across individuals and time. The SESOI is ideally based on a cost-benefit analysis. Since both costs and benefits are necessarily relative, the SESOI will depend on the researcher who designs the study. The goal of setting a SESOI is to clearly justify why designing a study that has a high probability of rejecting effects larger than a specified value contributes to our knowledge base. Researchers should not aim to determine a SESOI that is universally valid. The goal is to set a SESOI such that inferences based on it answer a meaningful question.

@@

#Adressing Trello cards

For the example where the results are inconclusive, explain how adding more data, will, assuming the true effect is 0, lead to a smaller confidence interval, and thus will lead to an informative test. However, with increasing data, the effect could also slightly increase, to above or at the SESOI level, and then the effect will be significant and not equivalent. With massive amounts of data, many results will be significant and equivalent.

Boundary hacking should go into the discussion.
